---
title: "Identify flu seasons using the 4 census regions"
output: html_document
---

```{r load-libraries, echo=F, message=F}
library(tidyverse)
library(magrittr)
library(patchwork)
library(RcppRoll)
library(readxl)
library(lubridate)
library(scales)
library(splines)
library(here)
```

```{r define-census-regions, echo=F}
ne_states <- c("Connecticut", "Maine", "Massachusetts", "New Hampshire", 
                 "Rhode Island", "Vermont", "New Jersey", "New York", "Pennsylvania")

mw_states <- c("Illinois", "Indiana", "Michigan", "Minnesota", "Ohio", 
               "Wisconsin","North Dakota", "South Dakota",
               "Iowa", "Kansas", "Missouri", "Nebraska")

s_states <- c("Delaware", "District of Columbia", "Maryland",  
                 "Virginia", "West Virginia",
              "Tennessee", "Kentucky", "Mississippi", "Alabama",
                       "Florida", "Georgia", "North Carolina", "South Carolina",
              "Arkansas", "Louisiana", "Oklahoma", "Texas")

w_states <- c("Colorado", "Montana",  "Utah", 
                 "Wyoming", "New Mexico",
              "Arizona", "California", "Hawaii", "Nevada",
              "Alaska", "Idaho", "Oregon", "Washington")
```

## Overview

Our goal is to identify the weeks defining the start and end dates of season 
circulating Influenza separately for each Census Region

## Data 

The original data I received from Pingsheng did not include data pre-2010, so 
I was missing 1995-2009. I went to CDC fluview 
(https://gis.cdc.gov/grasp/fluview/fluportaldashboard.html)
and was able to download additional
data pre 2010, however, these data are by HHS region or Census division (cannot 
download data for smaller geographic units like the state). So I downloaded the 
data for all HHS. The earliest year these data go back to is 1997. 

```{r import-data, echo = F, message=F}

# data from pre 2015
# combined across public health and clinical laboratories

dat_9715 <- read_csv(here("Data", "WHO_NREVSS_Combined_prior_to_2015_16_By-Census-division.csv"), 
                          skip = 1,
                          na = "X"
)

table(dat_9715$REGION) #9 census divisions regions
table(dat_9715$YEAR) #1997 t0 2015 imported 

names(dat_9715)
```

```{r define-census-region, echo = F, message=F}
dat_9715 %<>% mutate(census_division = case_when(REGION %in% c("Mid-Atlantic", "New England") ~ "Northeast",
                                                 REGION %in% c("East South Central", "South Atlantic",
                                                               "West South Central") ~ "South",
                                                 REGION %in% c("West North Central", "East North Central") ~ "Midwest",
                                                 REGION %in% c("Pacific", "Mountain") ~ "West"))

table(is.na(dat_9715$census_division))
table(dat_9715$census_division, dat_9715$REGION)
```

```{r tidy-data-and-make-subsets, echo=F}
dat_9715 %<>% mutate(states = case_when(census_division == "Northeast" ~ paste(ne_states, collapse = ', '),
                                        census_division == "South" ~ paste(s_states, collapse = ', '),
                                        census_division == "Midwest" ~ paste(mw_states, collapse = ', '),
                                        census_division == "West" ~ paste(w_states, collapse = ', ')
                                        ))

```

```{r}
dat_9715 %<>% rename(num_test = `TOTAL SPECIMENS`,
                     num_pos_A_h1n1 = `A (2009 H1N1)`,
                     num_pos_A_h1 = `A (H1)`,
                     num_pos_A_h3 = `A (H3)`,
                     num_pos_A_unknown = `A (Subtyping not Performed)`,
                     num_pos_A_unable = `A (Unable to Subtype)`,
                     num_pos_B = B,
                     num_pos_h3n2v = H3N2v,
                     percent_pos = `PERCENT POSITIVE`)

dat_9715 %<>% mutate(num_pos = num_pos_A_h1n1 + num_pos_A_h1 + 
                       num_pos_A_h3 + num_pos_A_unknown + 
                       num_pos_A_unable + num_pos_B + num_pos_h3n2v,
                     pp2 = num_pos/num_test)

# dt3 is the date (YYYY-MM-DD) for each calendar week 
dat_9715 %<>% 
  mutate(dt = ymd(paste0(YEAR, "-01-01")), #start by appending the year only
         dt2 = dt + dweeks(WEEK - 1), #then add in the week info
         dt3 = case_when(
           YEAR == 1997 ~ dt2 - ddays(3), # Jan 1 is W, go back to last Sun
           YEAR == 1998 ~ dt2 + ddays(3), # Jan 1 is Th, go fwd to next Sun
           YEAR == 1999 ~ dt2 + ddays(2), # jan 1 is Fr, go fwd to next Sun
           YEAR == 2000 ~ dt2 + ddays(1), #leap year, jan 1 sat
           YEAR == 2001 ~ dt2 - ddays(1),
           YEAR == 2002 ~ dt2 - ddays(2),
           YEAR == 2003 ~ dt2 - ddays(3),
           YEAR == 2004 ~ dt2 + ddays(3), #leap year
           YEAR == 2005 ~ dt2 + ddays(1),
           YEAR == 2006 ~ dt2 - ddays(0),
           YEAR == 2007 ~ dt2 - ddays(1),
           YEAR == 2008 ~ dt2 - ddays(2), #leap year
           YEAR == 2009 ~ dt2 + ddays(3),
           YEAR == 2010 ~ dt2 + ddays(2), 
           YEAR == 2011 ~ dt2 + ddays(1),
           YEAR == 2012 ~ dt2 - ddays(0), #leap year
           YEAR == 2013 ~ dt2 - ddays(2),
           YEAR == 2014 ~ dt2 - ddays(3),
           YEAR == 2015 ~ dt2 + ddays(3)
         )
  )

#See this info for how to figure out what is the first week of the year
#https://ndc.services.cdc.gov/wp-content/uploads/MMWR_Week_overview.pdf
#if Jan 1 is S, M, Tu, or W, then the first week of the year begins on the Sunday of that week
# If Jan 1 is Th, F, or S, then the first week of the year begins on the Sunday of the following week


dat_9715 %<>% mutate(no_tests = num_test == 0, 
                     test_lt_100 = num_test < 100)

dat_9715 %<>% arrange(YEAR, WEEK) %>% mutate(counter = row_number())
#calculating the dataset size expected based on number of weeks per year and the number of HHS and the number of weeks in beginning and ending year
#(52 in reg year*13 reg years*10 HHS)+(53 in leap year*4 leap years*10)+(13 begin year*10)+(39 last year*10)

#issue; there is less unique dt3 then there are combinations of WEEK*YEAR
#this is because there are years where there are 53 weeks coded for that year and it ends up generating the same dt3 for the last week of the year and the first week of the next year
#this is because the last week of one year and the first week of the next year are the same week! But somehow the CDC has different data for those eg look at counter =  141-149 and the proceeding rows

#add one fabricated test that is false (so add one to num test but nothing to num positive)
dat_9715 %<>% mutate(num_test_p1 = num_test + 1)
```

```{r}
dat2 <- dat_9715 %>% group_by(census_division, dt3) %>% 
  summarise(states = first(states),
            YEAR = first(YEAR),
            WEEK = first(WEEK),
            num_test = sum(num_test),
            num_pos = sum(num_pos)) %>%
  mutate(pp2 = num_pos/num_test,
         num_test_p1 = num_test + 1,
         no_tests = num_test == 0, 
         test_lt_100 = num_test < 100)
```

```{r}
table(is.na(dat2$pp2)) 
#206 are missing percent_pos info (0 tests with 0 positive results)
```
Summary of the number of tests by Census division -- want to make sure data isn't
too noisy for any division.

```{r}
#six number summary of the number of tests per HHS region over time
dat2 %>% 
  group_by(census_division) %>%
  summarise(min = min(num_test), 
            Q1 = quantile(num_test, 0.25),
            mean = mean(num_test),
            med = median(num_test),
            Q3 = quantile(num_test, 0.75), 
            max = max(num_test))
```

```{r}
dat2 %<>%
  ungroup() %>% 
  rowwise() %>%
  mutate(lb = ifelse(num_test == 0, 
                     0, 
                     binom.test(x = num_pos,
                                n = num_test, 
                                conf.level = 0.95)$conf.int[1]),
         ub = ifelse(num_test == 0, 
                     1, 
                     binom.test(x = num_pos,
                                n = num_test, 
                                conf.level = 0.95)$conf.int[2]),
         ub2 = case_when(ub > 0.65 ~ 0.65,
                         ub <= 0.65 ~ ub))
```

```{r hhs1}
division <- "Northeast"
```

`r division`

```{r}
assign(paste0(division, "_summary"), dat2 %>% filter(census_division == division))
```

Proportion of the weeks in a calendar year with no tests

```{r}
get(paste0(division, "_summary")) %>% 
  group_by(YEAR) %>%
  summarise(num_weeks = n(),
            num_weeks_no_tests = sum(no_tests)) %>%
  mutate(prop_no_tests = round((num_weeks_no_tests/num_weeks)*100, 1))
```


```{r}
a <- ggplot(data = get(paste0(division, "_summary")), aes(x = dt3, y = num_test)) + 
  geom_line() + 
  labs(y = "Number of tests", x = "Year", 
       title = paste("Number of clinical tests in ", division, " 1997+")) +
  theme_bw() + 
  scale_y_log10() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") 

b <- ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2, fill = "95% CI")) +
  geom_line(col = "#7a0177") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "Percentage tests that are positive", x = "Year") +
  theme_bw() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") + 
  scale_fill_manual(values = c("#bdbdbd")) +
  scale_y_continuous(limits = c(0, 0.65))

a + b + plot_layout(nrow = 2)
```

```{r}

assign(paste0(division, "_mod0"),
       glm(num_pos ~ ns(dt3, df = 160), 
           offset = log(num_test_p1), #18 seasons with 10 df per season = 180
           data = get(paste0(division, "_summary")), family = "poisson")
)

Northeast_summary$p1 <- predict(Northeast_mod0, type = "response")
Northeast_summary$p2 <- Northeast_summary$p1/Northeast_summary$num_test_p1

ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), lwd = 1, col = "#fd8d3c") + 
  geom_line(col = "#7a0177") +
  geom_hline(yintercept = 0.05) +
  labs(y = "Percentage tests that are positive", x = "Date") +
  theme_bw() + labs(title = paste0(division, ": Actual vs. smoothed (df = 160)")) +
  scale_x_date(date_breaks = "years", date_labels = "%y")
```

**Identify flu season start and end date for HHS region 1**

```{r, echo = F}
Northeast_summary %<>% 
  ungroup() %>% 
  arrange(dt3) %>%
  mutate(gte5 = case_when(p2 >= 0.05 ~ 1, 
                          p2 < 0.05 ~ 0),
         diff_lag1_pp = gte5 - lag(gte5),
         diff_lag2_pp = gte5 - lag(gte5, n = 2),
         runsum = RcppRoll::roll_sum(diff_lag2_pp, 2, align = "right", fill = NA),
         start_flu_season = case_when(runsum == 2 ~ 1,
                                      runsum != 2 ~ 0),
         end_flu_season = case_when(runsum == -2 ~ 1,
                                    runsum != -2 ~ 0))
```

```{r make-dataset-of-flu-dates, echo = F}
Northeast_flu_season_start <- Northeast_summary %>% filter(start_flu_season == 1) %>% select(dt3)
dim(Northeast_flu_season_start) #16

Northeast_flu_season_end <- Northeast_summary %>% filter(end_flu_season == 1) %>% select(dt3)
dim(Northeast_flu_season_end) #16

Northeast_flu_season_dates <- cbind(Northeast_flu_season_start,
                              Northeast_flu_season_end)

names(Northeast_flu_season_dates) <- c("flu_season_start", "flu_season_end")

```

**List of flu season dates**

```{r, echo=F}
Northeast_flu_season_dates %<>% 
mutate(last_end = lag(flu_season_end),
       season_length = flu_season_end - flu_season_start,
       time_since_last = flu_season_start - last_end)

Northeast_flu_season_dates
#write_csv(Northeast_flu_season_dates, file = here("Data", "Northeast-flu-seasons-9715.csv"))
```

```{r, echo=F}
ne_plot <- ggplot(dat = Northeast_summary, 
       aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2), alpha = 0.4) +
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), col = "#7a0177") + 
  geom_line(col = "#fd8d3c") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "% positive flu test", x = "Year") +
  theme_bw() + labs(title = "Northeast") +
  geom_vline(data = Northeast_flu_season_dates, aes(xintercept = flu_season_end), col = "red") + 
  geom_vline(data = Northeast_flu_season_dates, aes(xintercept = flu_season_start), col = "green") +
  scale_x_date(date_breaks = "years", date_labels = "%y")

ne_plot

# ggsave(plot = ne_plot, 
#        filename = here("Plots", "Northeast-seasons-9715_2023-03-10.png"),
#        width = 10, height = 2, device = "png")
```

```{r}
write_csv(Northeast_summary, file = here("Data", "Northeast-weekly-smoothed.csv"))
```

```{r midwest}
division <- "Midwest"
```

`r division`

```{r}
assign(paste0(division, "_summary"), dat2 %>% filter(census_division == division))
```

Proportion of the weeks in a calendar year with no tests

```{r}
get(paste0(division, "_summary")) %>% 
  group_by(YEAR) %>%
  summarise(num_weeks = n(),
            num_weeks_no_tests = sum(no_tests)) %>%
  mutate(prop_no_tests = round((num_weeks_no_tests/num_weeks)*100, 1))
```

```{r}
a <- ggplot(data = get(paste0(division, "_summary")), aes(x = dt3, y = num_test)) + 
  geom_line() + 
  labs(y = "Number of tests", x = "Year", 
       title = paste("Number of clinical tests in ", division, " 1997+")) +
  theme_bw() + 
  scale_y_log10() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") 

b <- ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2, fill = "95% CI")) +
  geom_line(col = "#7a0177") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "Percentage tests that are positive", x = "Year") +
  theme_bw() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") + 
  scale_fill_manual(values = c("#bdbdbd")) +
  scale_y_continuous(limits = c(0, 0.65))

a + b + plot_layout(nrow = 2)
```

```{r}
assign(paste0(division, "_mod0"),
       glm(num_pos ~ ns(dt3, df = 160), 
           offset = log(num_test_p1), #18 seasons with 10 df per season = 180
           data = get(paste0(division, "_summary")), family = "poisson")
)

Midwest_summary$p1 <- predict(Midwest_mod0, type = "response")
Midwest_summary$p2 <- Midwest_summary$p1/Midwest_summary$num_test_p1

ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), lwd = 1, col = "#fd8d3c") + 
  geom_line(col = "#7a0177") +
  geom_hline(yintercept = 0.05) +
  labs(y = "Percentage tests that are positive", x = "Date") +
  theme_bw() + labs(title = paste0(division, ": Actual vs. smoothed (df = 160)")) +
  scale_x_date(date_breaks = "years", date_labels = "%y")
```

**Identify flu season start and end date for HHS region 1**

```{r, echo = F}
Midwest_summary %<>% 
  ungroup() %>% 
  arrange(dt3) %>%
  mutate(gte5 = case_when(p2 >= 0.05 ~ 1, 
                          p2 < 0.05 ~ 0),
         diff_lag1_pp = gte5 - lag(gte5),
         diff_lag2_pp = gte5 - lag(gte5, n = 2),
         runsum = RcppRoll::roll_sum(diff_lag2_pp, 2, align = "right", fill = NA),
         start_flu_season = case_when(runsum == 2 ~ 1,
                                      runsum != 2 ~ 0),
         end_flu_season = case_when(runsum == -2 ~ 1,
                                    runsum != -2 ~ 0))
```

```{r, echo = F}
Midwest_flu_season_start <- Midwest_summary %>% filter(start_flu_season == 1) %>% select(dt3)
dim(Midwest_flu_season_start) #18

Midwest_flu_season_end <- Midwest_summary %>% filter(end_flu_season == 1) %>% select(dt3)
dim(Midwest_flu_season_end) #18

Midwest_flu_season_dates <- cbind(Midwest_flu_season_start,
                              Midwest_flu_season_end)

names(Midwest_flu_season_dates) <- c("flu_season_start", "flu_season_end")

```

**List of flu season dates**

```{r, echo=F}
Midwest_flu_season_dates %<>% 
mutate(last_end = lag(flu_season_end),
       season_length = flu_season_end - flu_season_start,
       time_since_last = flu_season_start - last_end)

Midwest_flu_season_dates
write_csv(Midwest_flu_season_dates, file = here("Data", "Midwest-flu-seasons-9715.csv"))
```

```{r, echo=F}
mw_plot <- ggplot(dat = Midwest_summary, 
       aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2), alpha = 0.4) +
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), col = "#7a0177") + 
  geom_line(col = "#fd8d3c") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "% positive flu tests", x = "Year") +
  theme_bw() + labs(title = "Midwest") +
  geom_vline(data = Midwest_flu_season_dates, aes(xintercept = flu_season_end), col = "red") + 
  geom_vline(data = Midwest_flu_season_dates, aes(xintercept = flu_season_start), col = "green") +
  scale_x_date(date_breaks = "years", date_labels = "%y")

mw_plot

ggsave(plot = mw_plot, 
       filename = here("Plots", "Midwest-seasons-9715_2023-03-10.png"),
       width = 10, height = 2, device = "png")
```



```{r south}
division <- "South"
```

`r division`

```{r}
assign(paste0(division, "_summary"), dat2 %>% filter(census_division == division))
```

Proportion of the weeks in a calendar year with no tests

```{r}
get(paste0(division, "_summary")) %>% 
  group_by(YEAR) %>%
  summarise(num_weeks = n(),
            num_weeks_no_tests = sum(no_tests)) %>%
  mutate(prop_no_tests = round((num_weeks_no_tests/num_weeks)*100, 1))
```

```{r}
a <- ggplot(data = get(paste0(division, "_summary")), aes(x = dt3, y = num_test)) + 
  geom_line() + 
  labs(y = "Number of tests", x = "Year", 
       title = paste("Number of clinical tests in ", division, " 1997+")) +
  theme_bw() + 
  scale_y_log10() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") 

b <- ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2, fill = "95% CI")) +
  geom_line(col = "#7a0177") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "Percentage tests that are positive", x = "Year") +
  theme_bw() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") + 
  scale_fill_manual(values = c("#bdbdbd")) +
  scale_y_continuous(limits = c(0, 0.65))

a + b + plot_layout(nrow = 2)
```

```{r}
assign(paste0(division, "_mod0"),
       glm(num_pos ~ ns(dt3, df = 160), 
           offset = log(num_test_p1), #18 seasons with 10 df per season = 180
           data = get(paste0(division, "_summary")), family = "poisson")
)

South_summary$p1 <- predict(South_mod0, type = "response")
South_summary$p2 <- South_summary$p1/South_summary$num_test_p1

ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), lwd = 1, col = "#fd8d3c") + 
  geom_line(col = "#7a0177") +
  geom_hline(yintercept = 0.05) +
  labs(y = "Percentage tests that are positive", x = "Date") +
  theme_bw() + labs(title = paste0(division, ": Actual vs. smoothed (df = 160)")) +
  scale_x_date(date_breaks = "years", date_labels = "%y")
```

**Identify flu season start and end date for HHS region 1**

```{r, echo = F}
South_summary %<>% 
  ungroup() %>% 
  arrange(dt3) %>%
  mutate(gte5 = case_when(p2 >= 0.05 ~ 1, 
                          p2 < 0.05 ~ 0),
         diff_lag1_pp = gte5 - lag(gte5),
         diff_lag2_pp = gte5 - lag(gte5, n = 2),
         runsum = RcppRoll::roll_sum(diff_lag2_pp, 2, align = "right", fill = NA),
         start_flu_season = case_when(runsum == 2 ~ 1,
                                      runsum != 2 ~ 0),
         end_flu_season = case_when(runsum == -2 ~ 1,
                                    runsum != -2 ~ 0))
```

```{r, echo = F}
South_flu_season_start <- South_summary %>% filter(start_flu_season == 1) %>% select(dt3)
dim(South_flu_season_start) #18

South_flu_season_end <- South_summary %>% filter(end_flu_season == 1) %>% select(dt3)
dim(South_flu_season_end) #18

South_flu_season_dates <- cbind(South_flu_season_start,
                              South_flu_season_end)

names(South_flu_season_dates) <- c("flu_season_start", "flu_season_end")

```

**List of flu season dates**

```{r, echo=F}
South_flu_season_dates %<>% 
mutate(last_end = lag(flu_season_end),
       season_length = flu_season_end - flu_season_start,
       time_since_last = flu_season_start - last_end)

South_flu_season_dates
write_csv(South_flu_season_dates, file = here("Data", "South-flu-seasons-9715.csv"))
```

```{r, echo=F}
south_plot <- ggplot(dat = South_summary, 
       aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2), alpha = 0.4) +
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), col = "#7a0177") + 
  geom_line(col = "#fd8d3c") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "% positive flu tests", x = "Year") +
  theme_bw() + labs(title = "South") +
  geom_vline(data = South_flu_season_dates, aes(xintercept = flu_season_end), col = "red") + 
  geom_vline(data = South_flu_season_dates, aes(xintercept = flu_season_start), col = "green") +
  scale_x_date(date_breaks = "years", date_labels = "%y")

south_plot

ggsave(plot = south_plot, 
       filename = here("Plots", "South-seasons-9715_2023-03-10.png"),
       width = 10, height = 2, device = "png")
```

```{r}
write_csv(South_summary, file = here("flu-season-data", "South-weekly-smoothed.csv"))
```


```{r west}
division <- "West"
```

`r division`

```{r}
assign(paste0(division, "_summary"), dat2 %>% filter(census_division == division))
```

Proportion of the weeks in a calendar year with no tests

```{r}
get(paste0(division, "_summary")) %>% 
  group_by(YEAR) %>%
  summarise(num_weeks = n(),
            num_weeks_no_tests = sum(no_tests)) %>%
  mutate(prop_no_tests = round((num_weeks_no_tests/num_weeks)*100, 1))
```

```{r}
a <- ggplot(data = get(paste0(division, "_summary")), aes(x = dt3, y = num_test)) + 
  geom_line() + 
  labs(y = "Number of tests", x = "Year", 
       title = paste("Number of clinical tests in ", division, " 1997+")) +
  theme_bw() + 
  scale_y_log10() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") 

b <- ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2, fill = "95% CI")) +
  geom_line(col = "#7a0177") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "Percentage tests that are positive", x = "Year") +
  theme_bw() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") + 
  scale_fill_manual(values = c("#bdbdbd")) +
  scale_y_continuous(limits = c(0, 0.65))

a + b + plot_layout(nrow = 2)
```

```{r}
assign(paste0(division, "_mod0"),
       glm(num_pos ~ ns(dt3, df = 160), 
           offset = log(num_test_p1), #18 seasons with 10 df per season = 180
           data = get(paste0(division, "_summary")), family = "poisson")
)

West_summary$p1 <- predict(West_mod0, type = "response")
West_summary$p2 <- West_summary$p1/West_summary$num_test_p1

ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), lwd = 1, col = "#fd8d3c") + 
  geom_line(col = "#7a0177") +
  geom_hline(yintercept = 0.05) +
  labs(y = "Percentage tests that are positive", x = "Date") +
  theme_bw() + labs(title = paste0(division, ": Actual vs. smoothed (df = 160)")) +
  scale_x_date(date_breaks = "years", date_labels = "%y")
```

**Identify flu season start and end date for HHS region 1**

```{r, echo = F}
West_summary %<>% 
  ungroup() %>% 
  arrange(dt3) %>%
  mutate(gte5 = case_when(p2 >= 0.05 ~ 1, 
                          p2 < 0.05 ~ 0),
         diff_lag1_pp = gte5 - lag(gte5),
         diff_lag2_pp = gte5 - lag(gte5, n = 2),
         runsum = RcppRoll::roll_sum(diff_lag2_pp, 2, align = "right", fill = NA),
         start_flu_season = case_when(runsum == 2 ~ 1,
                                      runsum != 2 ~ 0),
         end_flu_season = case_when(runsum == -2 ~ 1,
                                    runsum != -2 ~ 0))
```

```{r, echo = F}
West_flu_season_start <- West_summary %>% filter(start_flu_season == 1) %>% select(dt3)
dim(West_flu_season_start) #18

West_flu_season_end <- West_summary %>% filter(end_flu_season == 1) %>% select(dt3)
dim(West_flu_season_end) #18

West_flu_season_dates <- cbind(West_flu_season_start,
                              West_flu_season_end)

names(West_flu_season_dates) <- c("flu_season_start", "flu_season_end")

```

**List of flu season dates**

```{r, echo=F}
West_flu_season_dates %<>% 
mutate(last_end = lag(flu_season_end),
       season_length = flu_season_end - flu_season_start,
       time_since_last = flu_season_start - last_end)

West_flu_season_dates
write_csv(West_flu_season_dates, file = here("Data", "West-flu-seasons-9715.csv"))
```

```{r, echo=F}
west_plot <- ggplot(dat = West_summary, 
       aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2), alpha = 0.4) +
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), col = "#7a0177") + 
  geom_line(col = "#fd8d3c") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "% positive flu tests", x = "Year") +
  theme_bw() + labs(title = "West)") +
  geom_vline(data = West_flu_season_dates, aes(xintercept = flu_season_end), col = "red") + 
  geom_vline(data = West_flu_season_dates, aes(xintercept = flu_season_start), col = "green") +
  scale_x_date(date_breaks = "years", date_labels = "%y")

west_plot

ggsave(plot = west_plot, 
       filename = here("Plots", "West-seasons-9715_2023-03-10.png"),
       width = 10, height = 2, device = "png")
```
```{r}
all_CR_flu_season_dates <- 
bind_rows(Northeast_flu_season_dates %>% mutate(Census_region = "Northeast", 
                                           states = paste(ne_states, collapse = ', ')),
          Midwest_flu_season_dates %>% mutate(Census_region = "Midwest", 
                                           states = paste(mw_states, collapse = ', ')),
          South_flu_season_dates %>% mutate(Census_region = "South", 
                                           states = paste(s_states, collapse = ', ')),
          West_flu_season_dates %>% mutate(Census_region = "West", 
                                           states = paste(w_states, collapse = ', ')))


write_csv(all_CR_flu_season_dates, file = here("Data", "all_CR-flu-seasons-1997-2015.csv"))  


```



```{r archive, include=F}
# fam <- family(Northeast_mod0)
# str(fam)
# ilink <- fam$linkinv
# ilink
# 
# Northeast_summary <- bind_cols(Northeast_summary, setNames(as_tibble(predict(Northeast_mod0, se.fit = TRUE)[1:2]),
#                                           c('fit_link','se_link'))
#                    )
# 
# Northeast_summary %<>% 
#   mutate(p1 = predict(Northeast_mod0, type ="response"), #predicted number who tests positive
#          p2 = p1/num_test_p1, #predicted positivity % using the model
#          fit_resp  = ilink(fit_link),
#          upr = ilink(fit_link + (2 * se_link)),
#          lwr = ilink(fit_link - (2 * se_link)),
#          upr2 = ilink(fit_link + (2 * se_link))/num_test_p1,
#          lwr2 = ilink(fit_link - (2 * se_link))/num_test_p1,
#          upr3 = case_when(upr2 > 0.65 ~ 0.65, 
#                           upr2 <= 0.65 ~ upr2),
#          ci_width = upr2 - lwr2)

#there is something wrong with the CIs
#for ex in 98-99, the CI is large and then goes back to being precise even during 
#weeks when 0 tests and 0 testing positive. Does not make any sense to me why
#the confidence interval would go back down to zero. 
#there is something wrong with the code oe something else...
# ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
#   geom_ribbon(aes(ymin = lwr2, ymax = upr3), alpha = 0.4) +
#   scale_y_continuous(limits = c(-0.05, 0.65)) + 
#   geom_line(aes(y = p2), lwd = 1, col = "#fd8d3c") + 
#   geom_line(col = "#7a0177") +
#   geom_hline(yintercept = 0.05) +
#   labs(y = "Percentage tests that are positive", x = "Date") +
#   theme_bw() + labs(title = paste0(division, ": Actual vs. smoothed (df = 160)")) +
#   scale_x_date(date_breaks = "years", date_labels = "%y")
# 
# ggplot(dat = get(paste0(division, "_summary")) %>% filter(YEAR < 2003), aes(x = dt3, y = pp2)) +
#   #geom_line(aes(y = p2), lwd = 1, col = "#fd8d3c", lwd = 0.5) +
#   geom_point(aes(y= upr3), fill = "green", pch = 21) +
#   geom_line(aes(y= lwr2), col = "green", lwd = 0.5) +
#   scale_y_continuous(limits = c(-0.05, 0.65)) +
#   #geom_line(col = "#7a0177") +
#   geom_hline(yintercept = 0.05) +
#   labs(y = "Percentage tests that are positive", x = "Date") +
#   theme_bw() + labs(title = paste0(division, ": Actual vs. smoothed (df = 160)")) +
#   scale_x_date(date_breaks = "years", date_labels = "%y")
```

