---
title: "All HHS: Define flu seasons for 2015 to 2021 by Census Region"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r load-libraries, echo=F, message=F}
library(tidyverse)
library(magrittr)
library(patchwork)
library(RcppRoll)
library(readxl)
library(lubridate)
library(scales)
library(splines)
library(here)
```

```{r define-census-regions, echo=F}
ne_states <- c("Connecticut", "Maine", "Massachusetts", "New Hampshire", 
                 "Rhode Island", "Vermont", "New Jersey", "New York", "Pennsylvania")

mw_states <- c("Illinois", "Indiana", "Michigan", "Minnesota", "Ohio", 
               "Wisconsin","North Dakota", "South Dakota",
               "Iowa", "Kansas", "Missouri", "Nebraska")

s_states <- c("Delaware", "District of Columbia", "Maryland",  
                 "Virginia", "West Virginia",
              "Tennessee", "Kentucky", "Mississippi", "Alabama",
                       "Florida", "Georgia", "North Carolina", "South Carolina",
              "Arkansas", "Louisiana", "Oklahoma", "Texas")

w_states <- c("Colorado", "Montana",  "Utah", 
                 "Wyoming", "New Mexico",
              "Arizona", "California", "Hawaii", "Nevada",
              "Alaska", "Idaho", "Oregon", "Washington")
```

## Overview

Our goal is to identify the weeks defining the start and end dates of season 
circulating Influenza separately for each Census region. 

## Data

These data were sent by NREVSS to Pingsheng in response to a data request. A
couple files were sent and CR added a readme file to her copy of the data with 
more information. Briefly, for years 2015 and later, the public health and 
clinical lab data is delineated across two files. However,
the public health data is only for a state-season (rather than a state-week) and 
cannot be used to identify flu season start and end dates. This file starts with the
2015-2016 season and uses data as recent as week 40 of 2021.

```{r import-data, echo = F, message=F}

Clinical_2015 <- read_csv(here("Data", "raw", "WHO_NREVSS_Clinical_Labs_By-Census-division.csv"), 
                          skip = 1, 
                          na = "X")

PH_2015 <- read_csv(here("Data", "raw", "WHO_NREVSS_Public_Health_Labs_By-Census-division.csv"), 
                          skip = 1, 
                          na = "X")
names(Clinical_2015)
names(PH_2015)

Clinical_2015 %<>% filter(YEAR < 2022)
PH_2015 %<>% filter(YEAR < 2022)
```

```{r define-census-region, echo = F, message=F}
Clinical_2015 %<>% mutate(census_division = case_when(REGION %in% c("Mid-Atlantic", "New England") ~ "Northeast",
                                                      REGION %in% c("East South Central", "South Atlantic",
                                                                    "West South Central") ~ "South",
                                                      REGION %in% c("West North Central", "East North Central") ~ "Midwest",
                                                      REGION %in% c("Pacific", "Mountain") ~ "West"))

table(is.na(Clinical_2015$census_division))
table(Clinical_2015$census_division, Clinical_2015$REGION)

PH_2015 %<>% mutate(census_division = case_when(REGION %in% c("Mid-Atlantic", "New England") ~ "Northeast",
                                                REGION %in% c("East South Central", "South Atlantic",
                                                              "West South Central") ~ "South",
                                                REGION %in% c("West North Central", "East North Central") ~ "Midwest",
                                                REGION %in% c("Pacific", "Mountain") ~ "West"))

table(is.na(PH_2015$census_division))
table(PH_2015$census_division, PH_2015$REGION)
```

```{r tidy-data-and-make-subsets, echo=F}
Clinical_2015 %<>% mutate(states = case_when(census_division == "Northeast" ~ paste(ne_states, collapse = ', '),
                                        census_division == "South" ~ paste(s_states, collapse = ', '),
                                        census_division == "Midwest" ~ paste(mw_states, collapse = ', '),
                                        census_division == "West" ~ paste(w_states, collapse = ', ')
                                        ))

PH_2015 %<>% mutate(states = case_when(census_division == "Northeast" ~ paste(ne_states, collapse = ', '),
                                        census_division == "South" ~ paste(s_states, collapse = ', '),
                                        census_division == "Midwest" ~ paste(mw_states, collapse = ', '),
                                        census_division == "West" ~ paste(w_states, collapse = ', ')
                                        ))
```

```{r, echo=F}
Clinical_2015 %<>% rename(num_test = `TOTAL SPECIMENS`,
                          num_pos_A = `TOTAL A`,
                          num_pos_B = `TOTAL B`,
                          percent_pos = `PERCENT POSITIVE`,
                          percent_pos_A = `PERCENT A`,
                          percent_pos_B = `PERCENT B`)

Clinical_2015 %<>% mutate(num_pos = num_pos_A + num_pos_B)

# add Sunday date for each week of the year

Clinical_2015 %<>% 
  mutate(dt = ymd(paste0(YEAR, "-01-01")), #start by appending the year only
         dt2 = dt + dweeks(WEEK - 1), #then add in the week info
         dt3 = case_when(YEAR == 2015 ~ dt2 + ddays(3), #Sunday is a Th, add days to get to first week 
                         YEAR == 2016 ~ dt2 + ddays(2),
                         YEAR == 2017 ~ dt2 - ddays(0),
                         YEAR == 2018 ~ dt2 - ddays(1),
                         YEAR == 2019 ~ dt2 - ddays(2),
                         YEAR == 2020 ~ dt2 - ddays(3),
                         YEAR == 2021 ~ dt2 + ddays(2)
                         )
         )

#See this info for how to figure out what is the first week of the year
#https://ndc.services.cdc.gov/wp-content/uploads/MMWR_Week_overview.pdf
#if Jan 1 is S, M, Tu, or W, then the first week of the year begins on the sunday pf that week
# If Jan 1 is Th, F, or S, then the first week of the year begins on the sunday of the following week

#str(Clinical_2015$dt3)

PH_2015 %<>% rename(num_test = `TOTAL SPECIMENS`) %>%
  mutate(num_pos = `A (2009 H1N1)` + `A (H3)` + `A (Subtyping not Performed)` + 
           B + BVic + BYam + H3N2v,
         percent_pos = (num_pos/num_test)*100)

# add Sunday date for each week of the year
PH_2015 %<>% 
  mutate(dt = ymd(paste0(YEAR, "-01-01")), #start by appending the year only
         dt2 = dt + dweeks(WEEK - 1), #then add in the week info
         dt3 = case_when(YEAR == 2015 ~ dt2 + ddays(3), #Sunday is a Th, add days to get to first week 
                         YEAR == 2016 ~ dt2 + ddays(2),
                         YEAR == 2017 ~ dt2 - ddays(0),
                         YEAR == 2018 ~ dt2 - ddays(1),
                         YEAR == 2019 ~ dt2 - ddays(2),
                         YEAR == 2020 ~ dt2 - ddays(3),
                         YEAR == 2021 ~ dt2 + ddays(2)
                         )
         )
```


```{r}
Clinical_2015 %<>% mutate(no_tests = num_test == 0, 
                     test_lt_100 = num_test < 100)

Clinical_2015 %<>% arrange(YEAR, WEEK) %>% mutate(counter = row_number())

Clinical_2015 %<>% mutate(num_test_p1 = num_test + 1)

PH_2015 %<>% mutate(no_tests = num_test == 0, 
                     test_lt_100 = num_test < 100)

PH_2015 %<>% arrange(YEAR, WEEK) %>% mutate(counter = row_number())

PH_2015 %<>% mutate(num_test_p1 = num_test + 1)

```

```{r}
C2 <- Clinical_2015 %>% group_by(census_division, dt3) %>% 
  summarise(states = first(states),
            YEAR = first(YEAR),
            WEEK = first(WEEK),
            num_test = sum(num_test),
            num_pos = sum(num_pos)) %>%
  mutate(pp2 = num_pos/num_test,
         num_test_p1 = num_test + 1,
         no_tests = num_test == 0, 
         test_lt_100 = num_test < 100)

P2 <- PH_2015 %>% group_by(census_division, dt3) %>% 
  summarise(states = first(states),
            YEAR = first(YEAR),
            WEEK = first(WEEK),
            num_test = sum(num_test),
            num_pos = sum(num_pos)) %>%
  mutate(pp2 = num_pos/num_test,
         num_test_p1 = num_test + 1,
         no_tests = num_test == 0, 
         test_lt_100 = num_test < 100)
```

```{r}
table(is.na(C2$pp2)) 
table(is.na(P2$pp2)) 
#none are missing percent_pos info (0 tests with 0 positive results)
```

Summary of the number of tests by Census region -- want to make sure data isn't
too noisy for any division.

```{r}
#six number summary of the number of tests per HHS region over time
C2 %>% 
  group_by(census_division) %>%
  summarise(min = min(num_test), 
            Q1 = quantile(num_test, 0.25),
            mean = mean(num_test),
            med = median(num_test),
            Q3 = quantile(num_test, 0.75), 
            max = max(num_test))

P2 %>% 
  group_by(census_division) %>%
  summarise(min = min(num_test), 
            Q1 = quantile(num_test, 0.25),
            mean = mean(num_test),
            med = median(num_test),
            Q3 = quantile(num_test, 0.75), 
            max = max(num_test))
```

```{r}
C2 %<>% ungroup() %>% rowwise() %>%
  mutate(lb = ifelse(num_test == 0, 
                0, 
                binom.test(x = num_pos,
                           n = num_test, 
                           conf.level = 0.95)$conf.int[1]),
         ub = ifelse(num_test == 0, 
                1, 
                binom.test(x = num_pos,
                           n = num_test, 
                           conf.level = 0.95)$conf.int[2]),
         ub2 = case_when(ub > 0.65 ~ 0.65,
                         ub <= 0.65 ~ ub))

P2 %<>% ungroup() %>% rowwise() %>%
  mutate(lb = ifelse(num_test == 0, 
                0, 
                binom.test(x = num_pos,
                           n = num_test, 
                           conf.level = 0.95)$conf.int[1]),
         ub = ifelse(num_test == 0, 
                1, 
                binom.test(x = num_pos,
                           n = num_test, 
                           conf.level = 0.95)$conf.int[2]),
         ub2 = case_when(ub > 0.65 ~ 0.65,
                         ub <= 0.65 ~ ub))
```

```{r northeast}
division <- "Northeast"
```

`r division`

```{r}
assign(paste0(division, "_summary"), C2 %>% filter(census_division == division))

assign(paste0(division, "_summary2"), P2 %>% filter(census_division == division))
```

Proportion of the weeks in a calendar year with no tests

```{r}
get(paste0(division, "_summary")) %>% 
  group_by(YEAR) %>%
  summarise(num_weeks = n(),
            num_weeks_no_tests = sum(no_tests)) %>%
  mutate(prop_no_tests = round((num_weeks_no_tests/num_weeks)*100, 1))
```

```{r}
a <- ggplot(data = get(paste0(division, "_summary")), aes(x = dt3, y = num_test)) + 
  geom_line() + 
  labs(y = "Number of tests", x = "Year", 
       title = paste("Number of clinical tests in ", division, " 2015+")) +
  theme_bw() + 
  scale_y_log10() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") 

b <- ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2, fill = "95% CI")) +
  geom_line(col = "#7a0177") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "Percentage tests that are positive", x = "Year") +
  theme_bw() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") + 
  scale_fill_manual(values = c("#bdbdbd")) +
  scale_y_continuous(limits = c(0, 0.65))

a + b + plot_layout(nrow = 2)
```

```{r}
assign(paste0(division, "_mod0"),
       glm(num_pos ~ ns(dt3, df = 72), 
           offset = log(num_test),
           data = get(paste0(division, "_summary")), family = "poisson")
)

Northeast_summary$p1 <- predict(Northeast_mod0, type = "response")
Northeast_summary$p2 <- Northeast_summary$p1/Northeast_summary$num_test

ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), lwd = 1, col = "#fd8d3c") + 
  geom_line(col = "#7a0177") +
  geom_hline(yintercept = 0.05) +
  labs(y = "Percentage tests that are positive", x = "Date") +
  theme_bw() + labs(title = paste0(division, ": Actual vs. smoothed (df = 72)")) +
  scale_x_date(date_breaks = "years", date_labels = "%y")
```

**Identify flu season start and end date for HHS region 1**

```{r, echo = F}
Northeast_summary %<>% 
  ungroup() %>% 
  arrange(dt3) %>%
  mutate(gte5 = case_when(p2 >= 0.05 ~ 1, 
                          p2 < 0.05 ~ 0),
         diff_lag1_pp = gte5 - lag(gte5),
         diff_lag2_pp = gte5 - lag(gte5, n = 2),
         runsum = RcppRoll::roll_sum(diff_lag2_pp, 2, align = "right", fill = NA),
         start_flu_season = case_when(runsum == 2 ~ 1,
                                      runsum != 2 ~ 0),
         end_flu_season = case_when(runsum == -2 ~ 1,
                                    runsum != -2 ~ 0))
```

```{r, echo = F}
Northeast_flu_season_start <- Northeast_summary %>% filter(start_flu_season == 1) %>% select(dt3)
dim(Northeast_flu_season_start) #6

Northeast_flu_season_end <- Northeast_summary %>% filter(end_flu_season == 1) %>% select(dt3)
dim(Northeast_flu_season_end) #5

Northeast_flu_season_dates <- cbind(Northeast_flu_season_start,
                                    rbind(Northeast_flu_season_end, NA))

names(Northeast_flu_season_dates) <- c("flu_season_start", "flu_season_end")

Northeast_flu_season_dates %<>% filter(! year(flu_season_start) == 2021)

```

**List of flu season dates**

```{r, echo=F}
Northeast_flu_season_dates %<>% 
mutate(last_end = lag(flu_season_end),
       season_length = flu_season_end - flu_season_start,
       time_since_last = flu_season_start - last_end)

Northeast_flu_season_dates
#write_csv(Northeast_flu_season_dates, file = here("Data", "Northeast-flu-seasons-1521.csv"))
```

```{r, echo=F}
ne_plot2 <- ggplot(dat = Northeast_summary, 
       aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2), alpha = 0.4) +
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), col = "#7a0177") + 
  geom_line(col = "#fd8d3c") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "% positive flu tests", x = "Year") +
  theme_bw() + labs(title = "Northeast") +
  geom_vline(data = Northeast_flu_season_dates, aes(xintercept = flu_season_end), col = "red") + 
  geom_vline(data = Northeast_flu_season_dates, aes(xintercept = flu_season_start), col = "green") +
  scale_x_date(date_breaks = "years", date_labels = "%y")

ne_plot2

# ggsave(plot = ne_plot2, 
#        filename = here("Plots", "Northeast-seasons-1521_2023-03-10.png"),
#        width = 5, height = 2, device = "png")
```

```{r}
write_csv(Northeast_summary, file = here("Data", "Northeast-weekly-smoothed-2.csv"))
```

```{r midwest}
division <- "Midwest"
```

`r division`

```{r}
assign(paste0(division, "_summary"), C2 %>% filter(census_division == division))

assign(paste0(division, "_summary2"), P2 %>% filter(census_division == division))
```

Proportion of the weeks in a calendar year with no tests

```{r}
get(paste0(division, "_summary")) %>% 
  group_by(YEAR) %>%
  summarise(num_weeks = n(),
            num_weeks_no_tests = sum(no_tests)) %>%
  mutate(prop_no_tests = round((num_weeks_no_tests/num_weeks)*100, 1))
```

```{r}
a <- ggplot(data = get(paste0(division, "_summary")), aes(x = dt3, y = num_test)) + 
  geom_line() + 
  labs(y = "Number of tests", x = "Year", 
       title = paste("Number of clinical tests in ", division, " 2015+")) +
  theme_bw() + 
  scale_y_log10() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") 

b <- ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2, fill = "95% CI")) +
  geom_line(col = "#7a0177") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "Percentage tests that are positive", x = "Year") +
  theme_bw() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") + 
  scale_fill_manual(values = c("#bdbdbd")) +
  scale_y_continuous(limits = c(0, 0.65))

a + b + plot_layout(nrow = 2)
```

```{r}
assign(paste0(division, "_mod0"),
       glm(num_pos ~ ns(dt3, df = 72), 
           offset = log(num_test),
           data = get(paste0(division, "_summary")), family = "poisson")
)

Midwest_summary$p1 <- predict(Midwest_mod0, type = "response")
Midwest_summary$p2 <- Midwest_summary$p1/Midwest_summary$num_test

ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), lwd = 1, col = "#fd8d3c") + 
  geom_line(col = "#7a0177") +
  geom_hline(yintercept = 0.05) +
  labs(y = "Percentage tests that are positive", x = "Date") +
  theme_bw() + labs(title = paste0(division, ": Actual vs. smoothed (df = 72)")) +
  scale_x_date(date_breaks = "years", date_labels = "%y")
```

**Identify flu season start and end date for Midwest region 1**

```{r, echo = F}
Midwest_summary %<>% 
  ungroup() %>% 
  arrange(dt3) %>%
  mutate(gte5 = case_when(p2 >= 0.05 ~ 1, 
                          p2 < 0.05 ~ 0),
         diff_lag1_pp = gte5 - lag(gte5),
         diff_lag2_pp = gte5 - lag(gte5, n = 2),
         runsum = RcppRoll::roll_sum(diff_lag2_pp, 2, align = "right", fill = NA),
         start_flu_season = case_when(runsum == 2 ~ 1,
                                      runsum != 2 ~ 0),
         end_flu_season = case_when(runsum == -2 ~ 1,
                                    runsum != -2 ~ 0))
```

```{r make-dataset-of-flu-dates, echo = F}
Midwest_flu_season_start <- Midwest_summary %>% filter(start_flu_season == 1) %>% select(dt3)
dim(Midwest_flu_season_start) #6

Midwest_flu_season_end <- Midwest_summary %>% filter(end_flu_season == 1) %>% select(dt3)
dim(Midwest_flu_season_end) #5

Midwest_flu_season_dates <- cbind(Midwest_flu_season_start,
                                    rbind(Midwest_flu_season_end, NA))

names(Midwest_flu_season_dates) <- c("flu_season_start", "flu_season_end")

#remove the last "season" since don't have the 2022 data so don't want to capture this as ending
Midwest_flu_season_dates %<>% filter(! year(flu_season_start) == 2021)
```

**List of flu season dates**

```{r, echo=F}
Midwest_flu_season_dates %<>% 
mutate(last_end = lag(flu_season_end),
       season_length = flu_season_end - flu_season_start,
       time_since_last = flu_season_start - last_end)

Midwest_flu_season_dates
write_csv(Midwest_flu_season_dates, file = here("Data", "Midwest-flu-seasons-1521.csv"))
```

```{r, echo=F}
mw_plot2 <- ggplot(dat = Midwest_summary, 
       aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2), alpha = 0.4) +
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), col = "#7a0177") + 
  geom_line(col = "#fd8d3c") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "% positive flu tests", x = "Year") +
  theme_bw() + labs(title = "Midwest") +
  geom_vline(data = Midwest_flu_season_dates, aes(xintercept = flu_season_end), col = "red") + 
  geom_vline(data = Midwest_flu_season_dates, aes(xintercept = flu_season_start), col = "green") +
  scale_x_date(date_breaks = "years", date_labels = "%y")

mw_plot2

# ggsave(plot = mw_plot2, 
#        filename = here("Plots", "Midwest-seasons-1521_2023-03-10.png"),
#        width = 5, height = 2, device = "png")
```

```{r}
write_csv(Midwest_summary, file = here("Data", "Midwest-weekly-smoothed-2.csv"))
```

```{r south}
division <- "South"
```

`r division`

```{r}
assign(paste0(division, "_summary"), C2 %>% filter(census_division == division))

assign(paste0(division, "_summary2"), P2 %>% filter(census_division == division))
```

Proportion of the weeks in a calendar year with no tests

```{r}
get(paste0(division, "_summary")) %>% 
  group_by(YEAR) %>%
  summarise(num_weeks = n(),
            num_weeks_no_tests = sum(no_tests)) %>%
  mutate(prop_no_tests = round((num_weeks_no_tests/num_weeks)*100, 1))
```

```{r}
a <- ggplot(data = get(paste0(division, "_summary")), aes(x = dt3, y = num_test)) + 
  geom_line() + 
  labs(y = "Number of tests", x = "Year", 
       title = paste("Number of clinical tests in ", division, " 2015+")) +
  theme_bw() + 
  scale_y_log10() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") 

b <- ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2, fill = "95% CI")) +
  geom_line(col = "#7a0177") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "Percentage tests that are positive", x = "Year") +
  theme_bw() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") + 
  scale_fill_manual(values = c("#bdbdbd")) +
  scale_y_continuous(limits = c(0, 0.65))

a + b + plot_layout(nrow = 2)
```

```{r}
assign(paste0(division, "_mod0"),
       glm(num_pos ~ ns(dt3, df = 72), 
           offset = log(num_test),
           data = get(paste0(division, "_summary")), family = "poisson")
)

South_summary$p1 <- predict(South_mod0, type = "response")
South_summary$p2 <- South_summary$p1/South_summary$num_test

ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), lwd = 1, col = "#fd8d3c") + 
  geom_line(col = "#7a0177") +
  geom_hline(yintercept = 0.05) +
  labs(y = "Percentage tests that are positive", x = "Date") +
  theme_bw() + labs(title = paste0(division, ": Actual vs. smoothed (df = 72)")) +
  scale_x_date(date_breaks = "years", date_labels = "%y")
```

**Identify flu season start and end date for South region 1**

```{r, echo = F}
South_summary %<>% 
  ungroup() %>% 
  arrange(dt3) %>%
  mutate(gte5 = case_when(p2 >= 0.05 ~ 1, 
                          p2 < 0.05 ~ 0),
         diff_lag1_pp = gte5 - lag(gte5),
         diff_lag2_pp = gte5 - lag(gte5, n = 2),
         runsum = RcppRoll::roll_sum(diff_lag2_pp, 2, align = "right", fill = NA),
         start_flu_season = case_when(runsum == 2 ~ 1,
                                      runsum != 2 ~ 0),
         end_flu_season = case_when(runsum == -2 ~ 1,
                                    runsum != -2 ~ 0))
```

```{r, echo = F}
South_flu_season_start <- South_summary %>% filter(start_flu_season == 1) %>% select(dt3)
dim(South_flu_season_start) #7

South_flu_season_end <- South_summary %>% filter(end_flu_season == 1) %>% select(dt3)
dim(South_flu_season_end) #7

South_flu_season_dates <- cbind(South_flu_season_start,
                                    rbind(South_flu_season_end))

names(South_flu_season_dates) <- c("flu_season_start", "flu_season_end")

#remove the last "season" since don't have the 2022 data so don't want to capture this as ending
South_flu_season_dates %<>% filter(! year(flu_season_start) == 2021)
```

**List of flu season dates**

```{r, echo=F}
South_flu_season_dates %<>% 
mutate(last_end = lag(flu_season_end),
       season_length = flu_season_end - flu_season_start,
       time_since_last = flu_season_start - last_end)

South_flu_season_dates
write_csv(South_flu_season_dates, file = here("Data", "South-flu-seasons-1521.csv"))
```

```{r, echo=F}
south_plot2 <- ggplot(dat = South_summary, 
       aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2), alpha = 0.4) +
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), col = "#7a0177") + 
  geom_line(col = "#fd8d3c") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "% positive flu tests", x = "Year") +
  theme_bw() + labs(title = "South") +
  geom_vline(data = South_flu_season_dates, aes(xintercept = flu_season_end), col = "red") + 
  geom_vline(data = South_flu_season_dates, aes(xintercept = flu_season_start), col = "green") +
  scale_x_date(date_breaks = "years", date_labels = "%y")

south_plot2

ggsave(plot = south_plot2, 
       filename = here("Plots", "South-seasons-1521_2023-03-10.png"),
       width = 5, height = 2, device = "png")
```

```{r}
write_csv(South_summary, file = here("Data", "South-weekly-smoothed-2.csv"))
```

```{r west}
division <- "West"
```

`r division`

```{r}
assign(paste0(division, "_summary"), C2 %>% filter(census_division == division))

assign(paste0(division, "_summary2"), P2 %>% filter(census_division == division))
```

Proportion of the weeks in a calendar year with no tests

```{r}
get(paste0(division, "_summary")) %>% 
  group_by(YEAR) %>%
  summarise(num_weeks = n(),
            num_weeks_no_tests = sum(no_tests)) %>%
  mutate(prop_no_tests = round((num_weeks_no_tests/num_weeks)*100, 1))
```

```{r}
a <- ggplot(data = get(paste0(division, "_summary")), aes(x = dt3, y = num_test)) + 
  geom_line() + 
  labs(y = "Number of tests", x = "Year", 
       title = paste("Number of clinical tests in ", division, " 2015+")) +
  theme_bw() + 
  scale_y_log10() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") 

b <- ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2, fill = "95% CI")) +
  geom_line(col = "#7a0177") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "Percentage tests that are positive", x = "Year") +
  theme_bw() + 
  scale_x_date(date_breaks = "years", date_labels = "%y") + 
  scale_fill_manual(values = c("#bdbdbd")) +
  scale_y_continuous(limits = c(0, 0.65))

a + b + plot_layout(nrow = 2)
```

```{r}
assign(paste0(division, "_mod0"),
       glm(num_pos ~ ns(dt3, df = 72), 
           offset = log(num_test),
           data = get(paste0(division, "_summary")), family = "poisson")
)

West_summary$p1 <- predict(West_mod0, type = "response")
West_summary$p2 <- West_summary$p1/West_summary$num_test

ggplot(dat = get(paste0(division, "_summary")), aes(x = dt3, y = pp2)) + 
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), lwd = 1, col = "#fd8d3c") + 
  geom_line(col = "#7a0177") +
  geom_hline(yintercept = 0.05) +
  labs(y = "Percentage tests that are positive", x = "Date") +
  theme_bw() + labs(title = paste0(division, ": Actual vs. smoothed (df = 72)")) +
  scale_x_date(date_breaks = "years", date_labels = "%y")
```

**Identify flu season start and end date for West region 1**

```{r, echo = F}
West_summary %<>% 
  ungroup() %>% 
  arrange(dt3) %>%
  mutate(gte5 = case_when(p2 >= 0.05 ~ 1, 
                          p2 < 0.05 ~ 0),
         diff_lag1_pp = gte5 - lag(gte5),
         diff_lag2_pp = gte5 - lag(gte5, n = 2),
         runsum = RcppRoll::roll_sum(diff_lag2_pp, 2, align = "right", fill = NA),
         start_flu_season = case_when(runsum == 2 ~ 1,
                                      runsum != 2 ~ 0),
         end_flu_season = case_when(runsum == -2 ~ 1,
                                    runsum != -2 ~ 0))
```

```{r, echo = F}
West_flu_season_start <- West_summary %>% filter(start_flu_season == 1) %>% select(dt3)
dim(West_flu_season_start) #5

West_flu_season_end <- West_summary %>% filter(end_flu_season == 1) %>% select(dt3)
dim(West_flu_season_end) #5

West_flu_season_dates <- cbind(West_flu_season_start,
                                    rbind(West_flu_season_end))

names(West_flu_season_dates) <- c("flu_season_start", "flu_season_end")

#remove the last "season" since don't have the 2022 data so don't want to capture this as ending
West_flu_season_dates %<>% filter(! year(flu_season_start) == 2021)
```

**List of flu season dates**

```{r, echo=F}
West_flu_season_dates %<>% 
mutate(last_end = lag(flu_season_end),
       season_length = flu_season_end - flu_season_start,
       time_since_last = flu_season_start - last_end)

West_flu_season_dates
write_csv(West_flu_season_dates, file = here("Data", "West-flu-seasons-1521.csv"))
```

```{r, echo=F}
west_plot2 <- ggplot(dat = West_summary, 
       aes(x = dt3, y = pp2)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub2), alpha = 0.4) +
  scale_y_continuous(limits = c(-0.05, 0.65)) + 
  geom_line(aes(y = p2), col = "#7a0177") + 
  geom_line(col = "#fd8d3c") + 
  geom_hline(yintercept = 0.05) + 
  labs(y = "% positive flu tests", x = "Year") +
  theme_bw() + labs(title = "West") +
  geom_vline(data = West_flu_season_dates, aes(xintercept = flu_season_end), col = "red") + 
  geom_vline(data = West_flu_season_dates, aes(xintercept = flu_season_start), col = "green") +
  scale_x_date(date_breaks = "years", date_labels = "%y")

west_plot2

ggsave(plot = west_plot2, 
       filename = here("Plots", "West-seasons-1521_2023-03-10.png"),
       width = 5, height = 2, device = "png")
```

```{r}
write_csv(West_summary, file = here("Data", "West-weekly-smoothed-2.csv"))
```

```{r}
all_CR_flu_season_dates <- 
bind_rows(Northeast_flu_season_dates %>% mutate(Census_region = "Northeast", 
                                           states = paste(ne_states, collapse = ', ')),
          Midwest_flu_season_dates %>% mutate(Census_region = "Midwest", 
                                           states = paste(mw_states, collapse = ', ')),
          South_flu_season_dates %>% mutate(Census_region = "South", 
                                           states = paste(s_states, collapse = ', ')),
          West_flu_season_dates %>% mutate(Census_region = "West", 
                                           states = paste(w_states, collapse = ', ')))


write_csv(all_CR_flu_season_dates, file = here("Data", "all_CR-flu-seasons-2015-2021.csv"))  


```

The following chunk only runs if have plots loaded into memory from 
00_1997-2015_Identify-flu-seasons_Census-region.Rmd

```{r}
all_regions_and_year <- mw_plot + mw_plot2 + 
  ne_plot + ne_plot2 + 
  south_plot + south_plot2 + 
  west_plot + west_plot2 +
  plot_layout(nrow = 4, ncol = 2, widths = c(3, 1))

ggsave(plot = all_regions_and_year, 
       filename = here("Plots", "flu-seasons-for-all-regions-and-years.png"),
       width = 15, height = 10, device = "png", dpi = 300)
```

